BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2015-12-22 Paper2/BOAST.tex
Figures compressées...

\lstset{style=customc,
        caption={Laplace C99 Filter},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_c_filter}}
\begin{lstlisting}
void laplace(const int width,
             const int height,
             const unsigned char src[height][width][3],
                   unsigned char dst[height][width][3]){
  for (int j = 1; j < height-1; j++) {
    for (int i = 1; i < width-1; i++) {
      for (int c = 0; c < 3; c++) {
        int tmp = -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
                 - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
                 - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
        dst[j][i][c] = (tmp < 0 ? 0 : (tmp > 255 ? 255 : tmp));
      }
    }
  }
}
\end{lstlisting}

@@

  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
  uchar16 v12_ = vload16( 0, src + 3*width*(j-1) + 3*5*i     );
  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
  uchar16 v22_ = vload16( 0, src + 3*width*(j  ) + 3*5*i     );
  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
  uchar16 v32_ = vload16( 0, src + 3*width*(j+1) + 3*5*i     );
  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector_synth}}
\begin{lstlisting}
  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
  uchar16 v12_ = uchar16( v11_.s3456789a, v13_.s56789abc );
  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
  uchar16 v22_ = uchar16( v21_.s3456789a, v23_.s56789abc );
  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
  uchar16 v32_ = uchar16( v31_.s3456789a, v33_.s56789abc );
\end{lstlisting}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code}, 
        label={lst:BOAST_algebra},
        numbers=left }
\begin{lstlisting}
i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
k = BOAST::Int("k", :size => 8)
l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
BOAST::decl i, k, l
BOAST::pr i === 5
j = i + 5
BOAST::pr k === j * 2
BOAST::pr l[k] === 1.0
BOAST::register_funccall("sin")
BOAST::pr l[k+1] === BOAST::sin(j)
\end{lstlisting}

\begin{minipage}[b]{0.50\linewidth}
\centering
\lstset{style=customf,
        caption={Fortran output.},
        label={lst:BOAST_algebra_FORTRAN} }

\begin{lstlisting}
integer(kind=4) :: i
integer(kind=8) :: k
real(kind=8), dimension(-5:21) :: l
i = 5
k = (i + 5) * (2)
l(k) = 1.0_wp
l(k + 1) = sin(i + 5)
\end{lstlisting}
\end{minipage}
\hspace{0.04\linewidth}
\begin{minipage}[b]{0.44\linewidth}
\centering
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_algebra_C} }

\begin{lstlisting}
int32_t i;
int64_t k;
double l[27];
i = 5;
k = (i + 5) * (2);
l[k - (-5)] = 1.0;
l[k + 1 - (-5)] = sin(i + 5);
\end{lstlisting}
\end{minipage}
\caption{BOAST Code Snippet for Variables and Expressions.}
\label{fig:BOAST_algebra}
\end{figure}
The first and most fundamental abstraction is named Variable. Variables have a
type, a name, and a set of named attributes. The existing attributes are mainly
inspired from Fortran. Those attributes are not limited and can be arbitrarily
enriched, allowing a lot of flexibility in Variable management.

@@

\begin{lstlisting}
j = 0
do i = 0, 100, 1
  if (modulo(i, 7) == 0) then
    j = j + 1
  end if
end do
\end{lstlisting}
\end{minipage}
\hspace{0.04\linewidth}
\begin{minipage}[b]{0.47\linewidth}
\centering
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_control_C} }

\begin{lstlisting}
j = 0;
for (i = 0; i <= 100; i += 1) {
  if (modulo(i, 7) == 0) {
    j = j + 1;
  }
}
\end{lstlisting}
\end{minipage}
\caption{BOAST Code Snippet for control structures.}
\label{fig:BOAST_control}
\end{figure}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code}, 
        label={lst:BOAST_procedure},
        numbers=left }
\begin{lstlisting}
n = Int("n", :dir => :in)
ndat = Int("ndat", :dir => :in)
x = Real("x", :dir => :in, :dim => [ Dim(0, n-1), Dim(ndat) ])
y = Real("y", :dir => :out, :dim => [ Dim(ndat), Dim(0, n-1) ])
p = Procedure("magicfilter", [n, ndat, x, y])
opn p
close p
\end{lstlisting}

@@

\begin{lstlisting}
SUBROUTINE magicfilter(n, ndat, x, y)
  integer(kind=4), intent(in) :: n
  integer(kind=4), intent(in) :: ndat
  real(kind=8), intent(in), dimension(0:n - (1), ndat) :: x
  real(kind=8), intent(out), dimension(ndat, 0:n - (1)) :: y
END SUBROUTINE magicfilter
\end{lstlisting}
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_procedure_C} }

@@

\begin{table}
\centering
\begin{tabular}{c|r|r|c|r|c}
% \hline
  Image Size  & Naive (s) & Best (s) & Accel. & BOAST (s) & Accel. \\[4pt]
  \hline&&&&\\[-8pt]
768 x 432   & 0.0107    & 0.00669  & x1.6         & 0.000639  & x16.7 \\
% \hline
2560 x 1600 & 0.0850    & 0.0137   & x6.2         & 0.00687   & x12.4 \\
% \hline
2048 x 2048 & 0.0865    & 0.0149   & x5.8         & 0.00715   & x12.1 \\
% \hline
5760 x 3240 & 0.382     & 0.0449   & x8.5         & 0.0325    & x11.8 \\
% \hline
7680 x 4320 & 0.680     & 0.0747   & x9.1         & 0.0581    & x11.7 \\
% \hline
\end{tabular}
\caption{Best performance of ARM Laplace kernel.}
\label{tbl:ARM}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{c|r|r|c|r|c}
% \hline
  Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NV & Ratio \\[4pt]
  \hline&&&&\\[-8pt]
768 x 432   & 0.000639     & 0.000222    & x2.9  & 0.0000715    & x8.9  \\
% \hline
2560 x 1600 & 0.00687      & 0.00222     & x3.1  & 0.000782     & x8.8  \\
% \hline
2048 x 2048 & 0.00715      & 0.00226     & x3.2  & 0.000799     & x8.9  \\
% \hline
5760 x 3240 & 0.0325       & 0.0108      & x3.0  & 0.00351      & x9.3  \\
% \hline
7680 x 4320 & 0.0581       & 0.0192      & x3.0  & 0.00623      & x9.3  \\
% \hline
\end{tabular}
\caption{Best performance of Laplace Kernel on several architectures.}
\label{tbl:other}
\end{table}

>>>>

\lstset{style=customc,
        caption={Laplace C99 Filter},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_c_filter}}
\begin{lstlisting}
void laplace(const int width, const int height,
             const unsigned char src[height][width][3],
                   unsigned char dst[height][width][3]){
  for (int j = 1; j < height-1; j++) {
    for (int i = 1; i < width-1; i++) {
      for (int c = 0; c < 3; c++) {
        int tmp =
   -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
  - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
  - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
        dst[j][i][c] = (tmp<0 ? 0 : (tmp>255 ? 255 : tmp));
      }
    }
  }
}
\end{lstlisting}

@@

  uchar16 v11_ =vload16(0,src + 3*width*(j-1) + 3*5*i -3);
  uchar16 v12_ =vload16(0,src + 3*width*(j-1) + 3*5*i   );
  uchar16 v13_ =vload16(0,src + 3*width*(j-1) + 3*5*i +3);
  uchar16 v21_ =vload16(0,src + 3*width*(j  ) + 3*5*i -3);
  uchar16 v22_ =vload16(0,src + 3*width*(j  ) + 3*5*i   );
  uchar16 v23_ =vload16(0,src + 3*width*(j  ) + 3*5*i +3);
  uchar16 v31_ =vload16(0,src + 3*width*(j+1) + 3*5*i -3);
  uchar16 v32_ =vload16(0,src + 3*width*(j+1) + 3*5*i   );
  uchar16 v33_ =vload16(0,src + 3*width*(j+1) + 3*5*i +3);

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector_synth}}
\begin{lstlisting}
  uchar16 v11_ =vload16(0, src + 3*width*(j-1) + 3*5*i -3);
  uchar16 v13_ =vload16(0, src + 3*width*(j-1) + 3*5*i +3);
  uchar16 v12_ =uchar16(v11_.s3456789a, v13_.s56789abc);
  uchar16 v21_ =vload16(0, src + 3*width*(j  ) + 3*5*i -3);
  uchar16 v23_ =vload16(0, src + 3*width*(j  ) + 3*5*i +3);
  uchar16 v22_ =uchar16(v21_.s3456789a, v23_.s56789abc);
  uchar16 v31_ =vload16(0, src + 3*width*(j+1) + 3*5*i -3);
  uchar16 v33_ =vload16(0, src + 3*width*(j+1) + 3*5*i +3);
  uchar16 v32_ =uchar16(v31_.s3456789a, v33_.s56789abc);
\end{lstlisting}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code}, 
        label={lst:BOAST_algebra},
        numbers=right }
\begin{lstlisting}
i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
k = BOAST::Int("k", :size => 8)
l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
BOAST::decl i, k, l
BOAST::pr i === 5
j = i + 5
BOAST::pr k === j * 2
BOAST::pr l[k] === 1.0
BOAST::register_funccall("sin")
BOAST::pr l[k+1] === BOAST::sin(j)
\end{lstlisting}

\begin{minipage}[b]{0.5\linewidth}
\centering
\lstset{style=customf,
        caption={Fortran output.},
        label={lst:BOAST_algebra_FORTRAN} }

\begin{lstlisting}
integer(kind=4) :: i
integer(kind=8) :: k
real(kind=8), dimension(-5:21) :: l
i = 5
k = (i+5)*(2)
l(k) = 1.0_wp
l(k+1) = sin(i+5)
\end{lstlisting}
\end{minipage}
\hspace{0.08\linewidth}
\begin{minipage}[b]{0.40\linewidth}
\centering
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_algebra_C} }

\begin{lstlisting}
int32_t i;
int64_t k;
double l[27];
i = 5;
k = (i+5)*(2);
l[k-(-5)] = 1.0;
l[k+1-(-5)] = sin(i+5);
\end{lstlisting}
\end{minipage}
\caption{BOAST Code Snippet for Variables and Expressions.}
\label{fig:BOAST_algebra}
\end{figure}
The first and most fundamental abstraction is named Variable. Variables have a
type, a name, and a set of named attributes. The existing attributes are mainly
inspired from Fortran. Those attributes are not limited and can be arbitrarily
enriched, allowing a lot of flexibility in Variable management.

@@

\begin{lstlisting}
j=0
do i=0, 100, 1
  if(modulo(i, 7)==0) then
    j=j+1
  end if
end do
\end{lstlisting}
\end{minipage}
\hspace{0.08\linewidth}
\begin{minipage}[b]{0.40\linewidth}
\centering
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_control_C} }

\begin{lstlisting}
j=0;
for(i=0; i<=100; i+=1){
  if(modulo(i, 7)==0){
    j=j+1;
  }
}
\end{lstlisting}
\end{minipage}
\caption{BOAST Code Snippet for control structures.}
\label{fig:BOAST_control}
\end{figure}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code}, 
        label={lst:BOAST_procedure},
        numbers=right }
\begin{lstlisting}
n = Int("n", :dir=>:in)
ndat = Int("ndat", :dir=>:in)
x = Real("x", :dir=>:in, :dim=>[Dim(0, n-1), Dim(ndat)])
y = Real("y", :dir=>:out, :dim=>[Dim(ndat), Dim(0, n-1)])
p = Procedure("magicfilter", [n, ndat, x, y])
opn p
close p
\end{lstlisting}

@@

\begin{lstlisting}
SUBROUTINE magicfilter(n, ndat, x, y)
  integer(kind=4), intent(in) :: n
  integer(kind=4), intent(in) :: ndat
  real(kind=8), intent(in), dimension(0:n-(1), ndat) :: x
  real(kind=8), intent(out), dimension(ndat, 0:n-(1)) :: y
END SUBROUTINE magicfilter
\end{lstlisting}
\lstset{style=customc,
        caption={C output.},
        label={lst:BOAST_procedure_C} }

@@

\begin{table}
\caption{Best performance of ARM Laplace kernel.}
\label{tbl:ARM}
\centering
\begin{tabular}{c|r|r|c|r|c}
% \hline
  Image Size  & Naive (s) & Best (s) & Accel. & BOAST (s) & Accel. \\[4pt]
  \hline&&&&\\[-8pt]
768 x 432   & 0.0107    & 0.00669  & x1.6         & 0.000639  & x16.7 \\
% \hline
2560 x 1600 & 0.0850    & 0.0137   & x6.2         & 0.00687   & x12.4 \\
% \hline
2048 x 2048 & 0.0865    & 0.0149   & x5.8         & 0.00715   & x12.1 \\
% \hline
5760 x 3240 & 0.382     & 0.0449   & x8.5         & 0.0325    & x11.8 \\
% \hline
7680 x 4320 & 0.680     & 0.0747   & x9.1         & 0.0581    & x11.7 \\
% \hline
\end{tabular}
\end{table}

@@

\begin{table}
\caption{Best performance of Laplace Kernel on several architectures.}
\label{tbl:other}
\centering
\begin{tabular}{c|r|r|c|r|c}
% \hline
  Image Size & ARM & Intel & Ratio & NV & Ratio \\[4pt]
  \hline&&&&\\[-8pt]
768 x 432   & 0.000639     & 0.000222    & x2.9  & 0.0000715    & x8.9  \\
% \hline
2560 x 1600 & 0.00687      & 0.00222     & x3.1  & 0.000782     & x8.8  \\
% \hline
2048 x 2048 & 0.00715      & 0.00226     & x3.2  & 0.000799     & x8.9  \\
% \hline
5760 x 3240 & 0.0325       & 0.0108      & x3.0  & 0.00351      & x9.3  \\
% \hline
7680 x 4320 & 0.0581       & 0.0192      & x3.0  & 0.00623      & x9.3  \\
% \hline
\end{tabular}
\end{table}
